---
title: "FIN 411 Banking Analysis"
author: "Matt Brigida, Ph.D."
date: "December 3, 2019"
output:
  rmdformats::html_clean:
    highlight: kate
---

# RMarkdown

This is a markdown document in which code can be executed and output automatically included in the document.  The document can be rendered as HTML or pdf.

## An Example

The code chunk below will download the data we'll use in this analysis, and show the first 5 rows.

```{r cache = TRUE, message = FALSE, error = FALSE, warning = FALSE}
library(plm)

url <- "https://github.com/Matt-Brigida/FFIEC_Call_Reports/raw/master/1_querying_data_and_analysis/analyses/panel_data_analysis/full_panel/1_panel_with_full_quarter_date/1_one_panel_all_models/full_panel.rds"

panel <- readRDS(url(url))
```

```{r}
head(panel)
```

# Summary Statistics

First we'll create some tables of summary statistics to get a better understanding of our data set.  Questions we would like to answer are:

For now:

#. What is the capital structure of the average and median bank?
#. What is the average proportion of loans of a given type (consumer, C&I, real-estate)---as a percent of assets? 
#. Is bank capital structure different for young firms?
#. Do young firms focus on one type of loan?
#. Were capital structures different during/after the financial crisis?
#. What are the general characteristics of the average and median bank, i.e. ROA, NPAs

```{r results = 'asis', message = FALSE, error = FALSE, warning = FALSE}
library(stargazer)

var <- c("quarter", "totSBloans_Delt", "totNumSBloans_Delt", "t1_LR_lagged_1_year", "tot_SB_loans_TA_lagged_1", "ROA_lagged_1", "NPA_TA_lagged_1", "total_assets_lagged_1_year", "TD_TA_lagged_1", "de_novo", "TETA_lagged_1_year", "post_crisis_ind", "fin_crisis_ind")


panel_vars <- data.frame(panel[, var])

panel_vars <- panel_vars[complete.cases(panel_vars), ]


de_novos <- subset(panel_vars, de_novo == 1)

not_de_novos <- subset(panel_vars, de_novo == 0)

## create tables with stargazer

stargazer(de_novos[, -1], type = "html", title="Descriptive Statistics: De Novo Banks", digits=3, out="html", covariate.labels = c("% Change Amt. S. Bus. Loans", "% Change Num. S. Bus. Loans", "Tier 1 Leverage Ratio", "Small-Business Loans", "ROA", "NPA", "Total Assets", "Deposits", "De Novo", "Total Equity", "Post Crisis", "Financial Crisis"))


stargazer(not_de_novos[, -1], type = "html", title="Descriptive Statistics: Not De Novo Banks", digits=3, out="html", covariate.labels = c("% Change Amt. S. Bus. Loans", "% Change Num. S. Bus. Loans", "Tier 1 Leverage Ratio", "Small-Business Loans", "ROA", "NPA", "Total Assets", "Deposits", "De Novo", "Total Equity", "Post Crisis", "Financial Crisis"))
```


# Analyses

## The Change in Small Business Loans

```{r eval = TRUE, results = 'asis', message = FALSE, warning=FALSE, cache = TRUE}

panel <- panel[, var]

panel <- panel[complete.cases(panel), ]

## orthogonalize TE

TE_ortho <- lm(panel$TETA_lagged_1_year ~ panel$t1_LR_lagged_1_year)$resid

FEmodel1 <- plm(totSBloans_Delt ~  t1_LR_lagged_1_year + TE_ortho + tot_SB_loans_TA_lagged_1 + ROA_lagged_1 + NPA_TA_lagged_1 + I(log(panel$total_assets_lagged_1_year)) + TD_TA_lagged_1 + post_crisis_ind + fin_crisis_ind, data = panel, model = "within", effect = "individual")

## summary(FEmodel1)

FEmodel2 <- plm(totSBloans_Delt ~  t1_LR_lagged_1_year + TE_ortho + tot_SB_loans_TA_lagged_1 + ROA_lagged_1 + NPA_TA_lagged_1 + I(log(panel$total_assets_lagged_1_year))  + TD_TA_lagged_1 + post_crisis_ind + fin_crisis_ind + I(log(panel$total_assets_lagged_1_year) * ROA_lagged_1), data = panel, model = "within", effect = "individual")

## summary(FEmodel3)
## add de novo

stargazer(FEmodel1, FEmodel2, covariate.labels = c("T1LR", "TE", "Small Business Loans", "ROA", "NPA", "ln(TA)", "Deposits", "Post Crisis", "Fin Crisis", "ln(TA) * ROA", "ln(TA) * NPA"), dep.var.labels = "% Change in Amt. SB Loans", digits = 3, no.space=TRUE, header=FALSE, type='html', omit.stat=c("LL"), title = "All Banks: Determinants of the % Change in the Amount of Small-Business Loans", out = "html", intercept.bottom = TRUE, notes = "Results are from fixed-effects models with bank fixed effects, for the years 2001 through 2017.  Data are quarterly.  The dependent variable is percent change in the amount of small-business loans.  Small-Business loans are defined as the sum of commercial, industrial, and commercial real-estate loans.  All variables are lagged one year relative to the dependent variable.")
```

## Can We Predict A Financial Crisis?

```{r cache = TRUE}
library(pglm)

summary(pglm(fin_crisis_ind ~ NPA_TA_lagged_1 + ROA_lagged_1, data = panel, family = binomial('probit')))

```

# Graphics

```{r cache = TRUE}
library(highcharter)

chart_data <- subset(de_novos, quarter == 20050630)

hchart(de_novos, "bubble", hcaes(x = totSBloans_Delt, y = t1_LR_lagged_1_year, size = total_assets_lagged_1_year, color = NPA_TA_lagged_1)) %>%
    hc_add_theme(hc_theme_flatdark())

```
